### 2012-01-04 intro.tex
Intro

\section{Introduction}
\label{sec:intro}

% Talks about how AUV missions are traditionally planned. Why they might have
% been adequate in the past and are no longer so. How can deliberation
% (i.e. planning help) and how can it contribute to new ways of observing the
% ocean.