### 2012-01-04 related.tex
related work

\section{Related Work}
\label{sec:related}

% Refer to past work in AUV autonomy as well as robotic autonomy in
% general. This is general purpose and should be a good literature survey
% for robotics and AI.