### 2012-01-04 conclude.tex
updates

\section{Conclusion}

% Final parting words summarizing the contribution of deliberation towards
% vehicle autonomy.