### 2012-01-04 trunk/future.tex
update

\section{Future Work}
\label{sec:future}

% Where do we see all this going esp. in the context of embedded AI Planning
% for marine robots? This can be (and should be)
% speculative. Shore-side/onboard autonomy coordination and control of
% multiple vehicles is an important result of this activity.