### 2012-01-04 results.tex
update

\section{Experimental Results}
\label{sec:results}

% Show some of the results of the use of T-REX to convince people that is for
% real. CANON will be the predominant driver of the work.