### 2012-01-12 setobook.tex
updates

\documentclass[graybox,12pt]{svmult}

% \documentclass[12pt]{svmono}

\input{preamble}
\setcounter{secnumdepth}{-1}

% \usepackage[agsm]{harvard}

\title*{\textbf{\sc Deliberative Autonomy for Marine Robotics}}
\author{\normalsize Kanna~Rajan\inst{1} \and Fr\'ed\'eric Py\inst{1}
  \and Javier Barreiro\inst{2}}
\institute{Monterey Bay Aquarium Research Institute, California
  \texttt{\{Kanna.Rajan,fpy\}@mbari.org}
  \and
  NASA Ames Research Center, Moffett Field, California
  \texttt{javier.barreiro@nasa.gov}}
% \date{\small Monterey Bay Aquarium Research Institute \\
% 7700 Sandholdt Rd.\\
% Moss Landing,  CA, 95039\\
% {\tt }}

\begin{document}

\maketitle

\abstract{We describe a general purpose Artificial Intelligence based
  control architecture that incorporates in-situ decision making for
  adaptation and control for autonomous underwater vehicles
  (AUVs). The Teleo-Reactive EXecutive (\texttt{T-REX}) framework
  deliberates about future states, plans for actions and executes
  generated activities while monitoring plans for anomalous
  conditions. Plans are no longer scripted a priori but synthesized
  onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation. \rx is in routine operational use at MBARI,
  providing scientists a new tool to sample and observe the dynamic
  coastal ocean.}

\input{intro.tex}

\input{concepts.tex}

\input{related.tex}

\input{arch.tex}

\input{results.tex}

\input{future.tex}

\input{conclude.tex}

\section{Acknowledgements}

The authors are funded by a block grant from the David and Lucile
Packard Foundation to MBARI.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}