### 2012-01-18 trunk/future.tex
update

\section{Future Work}
\label{sec:future}

% Where do we see all this going esp. in the context of embedded AI Planning
% for marine robots? This can be (and should be)
% speculative. Shore-side/onboard autonomy coordination and control of
% multiple vehicles is an important result of this activity.

1.6 Future Work?
explain where EUROPA is going? (ANML, incorporation of classical planning algorithms, soft constraints, better search, etc).

The main advantage of CP is the richness of the constraints that can be modeled and the explicit use of logical inference to prune the search space