### 2012-01-19 trunk/related.tex
updates

\section{Background and Related Work}
\label{sec:related}

% Refer to past work in AUV autonomy as well as robotic autonomy in
% general. This is general purpose and should be a good literature survey
% for robotics and AI.

A distinct problem in Marine Robotics has been the use of the
overloaded term 'autonomy'. Not only does the notion transcend
different disciplines within engineering in this domain (e.g. the
Control Engineering community sees it distinctively from those in AI)
but users of marine platforms in oceanography as well as in maritime
defense conflate the methodology of use (tethered vs. untethered) with
their control. In this chapter, our notion of 'autonomy' deals
explicitly not only with the notion of dealing with the \emph{sense,
  plan and act} paradigm, but also the need to perform
\emph{computational search} between different action outcomes, an idea
central to the field of AI. Consequently, our literature survey here
is targeted towards a more focused and (to this chapter) relevant part
of the field of computation.

With few exceptions, most AUV control systems have been variations of
reactive approaches.  Typically, waypoint-based pre-defined mission
designs are uploaded to the AUV; specialized code fragments called
\emph{behaviors} are designed for the specific mission and a choice of
behaviors for the mission are used on the computational stack
\cite{bellingham94}. Swapping in and out of these behaviors using
conditionals or cleverer and quantitative approaches such as
\cite{Benjamin:2004} forms the basis for adaptation and safety in the
vehicle. Where such swapping does not aid adaptation, individual
behaviors are tweaked to generate some measure of responsiveness to
the surrounding environment \cite{yanwu08} to generate interesting
observations for oceanographers, yet technically naive in terms of
scalability and systematicity. \cite{Benjamin2006} is clear about the
extant of adaptation in stating:

{\footnotesize
  \begin{quote}
\small \emph{The primary difficulty often associated with behavior-based control
concerns action selection - namely how to ensure the chosen action
really is in the best overall interest of the robot or vehicle.}
\end{quote}
}

Such techniques have proved their mettle in the first round of use of
AUVs; they've provided operators a simple way to handle control
complexity of the vehicle while returned data at far higher resolution
than would have been possible using traditional ship-based methods and
at substantially reduced amortized costs of platform operation and
ownership over multiple years. AUV operators and clients for their
data products have reasons to be well satisfied with results thus
far. However, as hardware and more cost-effective sensors become more
robust to the environment and as mission durations increase (as
demonstrated by glider experiments such as \cite{rucool11}), sustained
presence in the ocean requires the ability to deal with off-nominal
conditions, that balances the needs of high-level mission management,
low-level navigation, instrument control and detection of unstructured
and poorly understood phenomena. Earlier reactive methods are unable
to perform such a balancing act without skewering the foundations of
the software development methodology leading to code bloat or worse.
\cite{carreras06} provides a reasonably good overview of AUV control
architectures that essentially is based on reactive control methods.

\begin{wrapfigure}{r}{0.2\textwidth}
 \centering
 \includegraphics[scale=0.3]{figs/planner-arch.pdf}
 \caption{\small A general architectural block diagram for an
   Artificial Intelligence based Planner.}
 \label{fig:planner}
\end{wrapfigure}

A word about generalized planning. Traditionally planning has been
considered computationally intensive (the general planning problem can
be worse than NP-complete \cite{ghallab04}). A major reason for this
belief had to do with the role of the planner (which was assumed to be
generative) and its place in the architecture (infrequently called to
re-plan the \emph{entire} mission in off-nominal conditions) within a
\emph{sense-plan-act} paradigm. This can limit the reactivity of a
robot especially when the environment could change at a rate faster
than the planner can plan.

The architecture of a typical AI planner is along the lines of
Fig. \ref{fig:planner}. Given a \emph{domain model} which encodes the
platform constraints in a higher-level language, an \emph{initial
  state}, a high-level objective or end \emph{goal}, the \emph{plan
  database} (or \texttt{plandb}) is where all assertions are tracked
and maintained. Theorem proving in the form of satisfaction of axioms
embodied in the domain model occurs within the \texttt{plandb}. It is
the job of the \emph{search engine} to provide the inferential
mechanism for state space exploration potentially aided by heuristics
encoded for \emph{search control}. In some complex domains
\cite{mus98} \emph{external modules} can provide additional domain
expertise which can augment the domain model. The end result is plan
formulation using a multitude of approaches either using forward or
backward search or using more generalized \emph{partial order} methods
\cite{ghallab04}.

Deliberative techniques for robotic control on the other hand
contrast in providing a distinct set of advantages:

\begin{enumerate}

\item action selection is driven axiomatically based on a systematic
  assessment of a range of causal conditions. An action is selected
  for instantiation only when there is causal support in the form of a
  constraint in a deterministic model of the vehicle's operation.

\item the system is goal directed which forms an objective towards
  which the system is expecting to converge providing a foci in action
  selection.

\item scaling to different (computational) problem size is dependent
  on incremental model building rather than being concerned about
  nuanced interactions between behaviors.

\item systematic software engineering methodologies such as spiral
  development \cite{boehm86} can further be used for scaling up the
  task as demonstrated by \cite{DS1report}.

\item when interactions between actions and/or behavior fragments do
  (and must) occur there are explicit constraints encoded in the model
  that must be computationally satisfied contrasting with carefully
  crafted strategies to instantiate behaviors on a stack. Such
  explicit recording of constraints in the model in turn aids
  long-term software maintenance and sustaining engineering.

\end{enumerate}

There are costs associated with such inference based control are also
not inconsequential. Foremost among them is the steep learning curve
that comes with the design of the domain model. With limited support
tools necessary for knowledge capture and design of the constraints
often in stylized higher-level languages (see \cite{NDDL} for an
example), it requires the modeler to think in alternative ways to do
problem solving. More importantly it is often necessary for the
modeler to be exposed to the internal representation of the planner
and how it performs search. THe level of expertise is often well above
what would be expected of a typical well-rounded Computer Science
graduate student.

However this is mild criticism given the end benefits and the
scientific and engineering goals associated with adaptive and
persistent observation for marine robots whether for civil or military
applications. Given the general problem-solving nature of such
deliberative systems, there is also a substantial overall benefit for
the field of robotics.

\subsection{Evolution of AI-based Robotic Planning: A Necessary Digression}

Planning and plan execution are not new to the overall field of
robotics. Early motivation of building computational mechanisms for
decision making were intended to be deployed on robotic devices by
very definition. The seminal volume \cite{computersthought} very
clearly articulates how physical manifestation of robots could be
controlled for task planning and execution. The very first planner
\cite{green69} was an exercise in computational theorem proving with
early state space planners popularized by \texttt{STRIPS}
\cite{strips71} dominating the academic landscape. An early
demonstration of planning as a situated agent was the highly
influential \texttt{Shakey} the robot experiment
\cite{shakey84}. Subsequent work driven largely by defense funding in
the United States sequestered the discipline in academic labs
principally applied for algorithmic development and if embedded then
in testing within the confines of modest indoor settings. The leap to
applying them for real world problem solving is relatively new and
driven largely by NASA applications pioneered by \cite{mus94,mus98,
  jonsson00, rajan00, chien05, bresina05}.

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale), notable examples of which include IPEM
\cite{AmbrosIngerson88}, ROGUE \cite{Haigh98}, the LAAS Architecture
\cite{alami:1998p820}, the Remote Agent Experiment \texttt{RAX}
\cite{mus98} and the Autonomous Spacecraft Experiment \texttt{ASE}
\cite{chien99} (see \cite{Knight01} for a survey). Early efforts were
cognizant of if absent with, dealing with time and resources for such
experiments.

Representationally, the first paper on managing time systematically
within AI Planning frameworks was \cite{dean87}; subsequent efforts by
Boddy and others \cite{Dean88,Boddy93} refined the notion of using a
temporal database and the conceptualization of temporal
intervals. Using these base-level concepts Muscettola and others
\cite{mus92, ghallab94, laborie95, cesta96} developed the notion of
plan/schedule envelopes using the notion of state variable
instantiated timelines. Simultaneously work by Dechter et.al
\cite{dechter91} resulted in efficient temporal constraint propagation
which systematically defined the notion of Simple Temporal Networks
(STNs) and path and arc consistency algorithms. This work neatly put
together earlier efforts by Mackworth and Freuder \cite{mackworth77,
  mack85} and the Constraints community on propagation algorithms. In
parallel work in the UK as a result of US DARPA funding, resulted in
the design of \texttt{O-Plan} \cite{currie91}, which leveraged the
notion of constraints and metric time, but within a Hierarchical Task
Network (HTN) representation. All these efforts were soon after Vere’s
paper on planning and time \cite{vere83} which had a profound effect
on the AI Planning community. In particular the work of Muscettola
\cite{mus94} was embraced by NASA for telescope scheduling. Around the
same time period, researchers at LAAS came up with an architecture
\cite{alami:1998p820} which encapsulated the \texttt{IxTeT} planner
\cite{ghallab94}. It featured advanced concepts for a temporal planner
with the notion of chronicles, plan recognition for partial plan
fragments and early use of systematic resource
constraints. Development on \texttt{IxTeT} has continued albeit at a
slower pace with the contribution of various PhD students’ thesis.

\subsection{The Remote Agent Experiment and beyond}

While \texttt{HSTS} \cite{mus94} was being implemented as a
ground-based planning tool for decision support for US based space
observations (EUVE and Cassini) \cite{mus95}, the opportunity to fly
onboard NASA’s New Millennium Deep Space $1$ came about. The design of
the Remote Agent Experiment \texttt{RAX} \cite{pell97, bernard98,
  pell98, mus98, DS1report, rajan00, jonsson00} was a direct off-shoot
of this effort where \texttt{HSTS} (written in LISP) was flown on
board and successfully commanded the spacecraft for a week in May
1999. There were a number of significant software engineering lessons
learned with the HSTS technology as integrated for the RAX:

\begin{enumerate}

\item Constraint-based representations were not only sufficient for
  plan synthesis, but valuable during debugging and development as a
  means of building a viable domain model.

\item Domain models when separated from the search engine as
  articulated by the model-based approach \cite{williams96a} ensured
  that sufficient effort would be targeted on human validation of the
  model building phase while ensuring that search engine stability
  across applications and domains resulted in lower cost to deploy.

\item Flexible temporal representation generating a range of plans as
  against a single plan, were robust for plan execution.

\item If planning and execution were disconnected (as in
  \texttt{RAX}), dispatchabilty \cite{mus98a} and controllability
  \cite{morris00} issues within temporal plans need be addressed. In
  \texttt{RAX}, a post-processing step was added to counter these
  effects. It was clear that execution needed access to the planners
  database and be able to propagate execution time constraints prior
  to command dispatching.

\end{enumerate}

It was with the demonstration of \texttt{RAX} $65$ Million miles in
outer space, that temporal reasoning using methods came to the
forefront of AI planning and for situated agents. Lessons learned from
\texttt{RAX} led to the development of \texttt{IDEA}
\cite{mus02,mus04} with the central theme of using planners
collectively for problem solving arose. Another critical theme was
iteratively and incremental plan repair of an anytime plan
\cite{Zaimag96} as proposed by \texttt{CASPER}
\cite{chien00}.

This last lesson in particular had a lasting impact with the
observation that planning and execution are strongly
inter-related. This core concept was behind the next generation of the
Remote Agent, called \texttt{IDEA} (Intelligent Deployable Execution
Agents) where planning and execution were \emph{intertwined} within a
single domain model that spanned the most abstract (for high-level
planning) to the least (for diagnosis). \texttt{IDEA} \cite{mus02,
  mus04} agents were expected to interact to achieve plan formulation;
however there was no systematic framework for formally governing these
interactions. \texttt{IDEA} was also computationally heavy and
required substantial effort to customize for an application domain. It
did move to a retooled version of \texttt{HSTS}, now called EUROPA
\cite{frank2003, barreiro09} and ported to the C++ language.

While \texttt{IDEA} was being deployed as a coordinating executive for
earth analog rover deployments \cite{wetter05} a separate development
was undertaken to deploy the EUROPA planner for NASA’s Mars
Exploration Rovers (MER) mission. The \texttt{MAPGEN}
(Mixed-initiative Activity Plan GENerator) planner \cite{bresina03,
  aichang04, bresina05, bresina05a} as a decision support tool in the
mixed-initiative mode; \texttt{MAPGEN} continues to be used to this
day to command the twin rovers on the Red
Planet. % While EUROPA2 was not
% designed as an embedded planner, NASA undertook a bottom-up
% re-implementation of the planner and substantially increased its
% performance. EUROPA2 was then released as open source to the AI
% community at large [Europa].

The Autonomous Sciencecraft Experiment (\texttt{ASE}) \cite{chien99,
  chien03} was another autonomy demonstration in space, this time for
a more observable vehicle, the EO-1 in Earth orbit. It encapsulated a
classical 3-layered architecture like \texttt{RAX}. The fundamental
functional difference was that goals could not only be sent by a
ground segment, but also by an onboard science driver which could
encapsulate a range of Machine Learned feature detectors which could
trigger the planner. However ASE was not an integrated system; the
\texttt{CASPER} planner \cite{chien00} would generate plans which were
executed by a separate commercially available rule-based system
\cite{icl}. \texttt{CASPER} touts a \emph{continuous} paradigm for
planning; however this incremental modification simply takes the
overall mission plan and selectively plans near-term at a more
detailed level. The system also uses iterative-repair as a way to fix
plans; however this too is derived from a model of the mission and
spacecraft constraints which allow simple reconfiguration modes
towards replanning. Robust software engineering is also not an
important aspect targeted by \texttt{ASE} given the disparate models
between the planner and the rule-based executive. With such a system,
real-time updates from the environment cannot be accommodated since
the planner is a monolithic computational environment. The rate of
updates from the science drivers is roughly in the order of an earth
orbit on the EO-1 vehicle. As a result, the ASE is only an incremental
update on RAX and is not suitable for domains where complex automated
reasoning and asynchronous environmental observations need to be
factored in deliberation.

\texttt{TCA} \cite{simmons94} is another framework for robot control
and has similar motivations to \texttt{RAX} and
\texttt{ASE}. Reactivity and deliberation are also key to \texttt{TCA}
and a systematic approach to development led to its design. However it
suffers from three key weaknesses. First it has a weaker
representational framework of Task Trees which require representation
of tasks within cleanly formulated hierarchies. Constraints between
branches within a task hierarchy are allowed; timing between leaves of
different trees is however not possible. Second, its temporal
framework does not deal with flexibility; time points are fixed and
while partial orders are possible, it does deal with execution time
uncertainty. Third, while \texttt{TCA} shares the notion of using
domain expertise of different planners, it does so thru a centralized
message passing mechanism. While this allows disparate code-bases to
talk to each other thru IPC (a inter-process communication or message
passing mechanism), failure in the centralized controller is
tantamount to a system crash.

The \texttt{ReSSAC} project at ONERA \cite{teichteil07} has similar
motivations in controlling aerial UAV platforms for autonomous Search
and Rescue. Like \texttt{TCA}, the project uses a supervisor to
control planning and at the same time decouples the deliberation and
reactive components from the lower level functional layer. This
deliberate separation is partly because of the use of optimizing MDP’s
(Markov Decision Processes \cite{mdp93}) for planning which uses
probabilistic state transitions to determine the next course of action
based on perceived state. To overcome the typical problem of state
space explosion with MDPs, a local heuristic is used in order to
generate reachable goal states incrementally. This work however does
not deal with metric time with a modestly simple model.

\texttt{CIRCA} \cite{musliner95} is another effort to bring decision
making for real-world problems with the augmented need to have
\emph{verifiable} controllers synthesized in-situ to ensure that
within explored states, there are no erroneous transitions to
dangerous states by asynchronously generating Test-Action-Pairs
(TAPs).  These are annotated production rules consisting of a set of
tests (or preconditions) and a set of actions to take if all the tests
return true. TAPs are synthesized and scheduled by \texttt{CIRCA} and
provide a viable way to deal with a dynamically changing
world. \texttt{CIRCA} however differentiates between reasoning about
time and reasoning in real-time with the implication that reasoning
necessarily requires substantial computation for state space
exploration which negates the real-time (and thus real-world)
impact.

A key and early concern that dominated planning was that of
scalability.  The planning cycle in many approaches was monolithic
often making fast reaction times impractical when necessary for
embedded robotic agents. Many of these systems (which were
$3$-layered) also utilized very different techniques for specifying
each layer in the architecture resulting in duplication of effort and
a diffusion of knowledge.  The work we highlight in this chapter
builds on the approach used by \texttt{IDEA} \cite{mus02, mus04} in
utilizing a collection of controllers, each interleaving planning and
execution within a common framework. \texttt{IDEA} however, provided
no support for conflict resolution between controllers, nor does it
provide an efficient algorithm for integrating current state within a
controller, relying instead on a possibly exponential planning
algorithm. Efficient synchronization of state in a partitioned
structure is fundamental to making the approach effective in practice.

Our system, \texttt{T-REX}, the focus of this chapter, is an agent
control structure formally defined as a composition of coordinated
control loops with a recurring sense-plan-act (SPA)
cycle. \texttt{T-REX} was designed after carefully evaluating lessons
learned from the IDEA and owes substantially to the notion of
partitioned problem solving from that effort. However this
partitioning is systematic and methodologically principled where we
manage the information flow within the partitioned structure to ensure
consistency in order to direct the flow of goals and observations in a
timely manner. The resulting control structure improves scalability
since many details of each controller can be encapsulated within a
single control loop.  Furthermore, partitioning increases robustness
since controller failure can be localized to enable graceful system
degradation, making this an effective \emph{divide-and-conquer}
approach to the overall control problem.

A recent variation of the key idea of such controller composition
derived from \texttt{T-REX} is \texttt{ROAR} \cite{degroote11} where
lower-level functional modules that control specific sensors are
organized within a graph hierarchy. One reason they cite such a
structure is essential is in dealing robustly to failure; an
off-nominal condition will allow rapid graph traversal to identify
alternative ways in which sensing tasks can be executed and therefore
aiding in responsiveness. However, it is unclear whether to date this
system actually deliberates and whether it is actually instantiated on
a real platform for control.

% ``The whole partition must be accordingly reorganized: this kind of
% construction does not scale well over a large variety of missions,
% missing the objective of a versatile architecture for robots.''