### 2012-02-15 concepts.tex
rearranging text from concepts into europa-basics

\subsubsection{\eu Plan Representation}
\label{sec:europa:pr}

This section describes the fundamental entities within \eu which are
typical to other CAIP-based formalisms.

\begin{description}

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time would be used.  As
  explained in the \textsf{CP} section, a variable can take values
  from a domain. In \eu, domains are defined over a specific data
  type, the primitive data types supported are:

  \begin{enumerate}
  \item \textit{String}: sequences of characters, like "Red",
    "Spirit", etc

  \item \textit{Boolean}: {True,False}

  \item \textit{Numeric}: integers and floats

  \item \textit{Object}: reference to an object instance (see below
    for a description of objects in \eu)

\end {enumerate}

Two types of variable Domains can be defined over a data type:

  \begin{enumerate}
  \item \textit{Enumerated}: Finite sets over any data type, specified
    explicitly, for instance: ["Red","Yellow","Green"], [1,18,32], etc
  \item \textit{Interval}: Only for Numeric data types, they are
    specified by a [LowerBound,UpperBound] pair, for instance: [1,10],
    [5.0,100.0], etc.
  \end {enumerate}

  \eu's ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans. \comment{say more about the benefits of flexibility?}

\item[\textbf{Constraints}] valid plans in most problem domains have
  to satisfy a number of restrictions, for instance, a Shopping Agent
  may only be able to perform one task at a time, may have limited
  time or an energy or fuel budget to perform a task, and so on. In
  \eu, these restrictions are represented by Constraints, which can be
  defined over any combination of Variables in a domain. \eu provides
  a Constraint Library that already implements many useful constraints
  (temporal, resource, relational, etc), it is also possible to add
  new constraints if required by a particular domain.

\item[\textbf{Objects}] The items we wish to describe and refer to in
  a domain are considered Objects. As in the case with object-oriented
  analysis and design, one can seek out the nouns in any domain
  description to find likely objects. In the Shopping Agent example,
  we might consider the Agent itself, Products and Locations all to be
  objects. Objects have state and behavior. For example, a Shopping
  Agent can have:

  \begin{itemize}

  \item State: its location, a bag that contains the products it has
    already purchased (the bag can in turn be another object), etc.

  \item Behavior: go to a location to look for a product, perform a
    purchase, return home, etc.

  \end{itemize}

  Objects that have similar state and behavior can be described
  generically in terms of object types or classes. \eu allows the
  definition of classes in the same way that it is done in popular
  object-oriented programming languages, where a class can have
  attributes (represented by Variables), and classes can be arranged
  in a single-parent hierarchy through inheritance.  However, one key
  aspect needs to be introduced to class definition in \eu that is not
  found in object-oriented programming: to describe state and behavior
  for the purposes of planning we need to build on the formalism of
  first order logic as explained below.

\item[\textbf{Tokens}] In first order logic, a predicate defines a
  relation between objects and properties \comment{citation?}.  In
  \eu, we define such relations between variables whose domains are
  sets of objects and sets of properties to describe state and
  behavior. For example, we might use a predicate \texttt{At($a$,$l$)}
  to indicate that agent $a$ is at location $l$, or a predicate
  \texttt{Buy($a$,$p$}) to indicate that agent $a$ is taking acting to
  buy product $p$. Note that $a$ is a variable which may have a number
  of possible values in a problem with multiple agents. Similarly, $l$
  and $p$ are variables whose values are the set of possible positions
  and products respectively. \eu can be used to create partial plans,
  where the domains of variables $a$ and $l$ can have more than one
  possible value in them, or grounded plans, where single values will
  be specified for each variable as we saw in the case of the
  N-queen's problem.

  In general, to come up with an executable plan it is not sufficient
  to state predicates that describe required state or behavior without
  also specifying some temporal extent over which each of those
  predicates hold. Predicates that are always true can be thought to
  hold from the beginning to the end of time. However, in practice,
  the temporal extent of interest must be defined with timepoints to
  represent its start and end. So we might want to write
  \texttt{At($a$,$s$,$e$,$l$)} to indicate that the agent $a$ is at
  location $p$ from time $s$ to time $e$. In fact, this pattern of
  using such predicates to describe both state and behavior of objects
  is so prevalent in \eu that we have introduced a special construct
  called a Token which has the built in variables to indicate the
  object to which the statement principally applies and the timepoints
  over which it holds. A Token is an instance of a predicate that
  represents and object's state or behavior and is defined over a
  temporal extent. In \eu, all predicate instances are Tokens. Also,
  in the same way that objects are described generically by classes,
  in \eu Tokens can be described generically by Token Types. Every
  token has five built-in variables:

  \begin{enumerate}

  \item \textit {start}: The beginning of the temporal extent over
    which the predicate is defined.

  \item \textit {end}: The end of the temporal extent over which the
    predicate is defined.

  \item \textit {duration}: The duration of the temporal extent over
    which the predicate is defined. The constraint \textit{start} $+$
    \textit{duration} = \textit{end} is enforced automatically.

  \item \textit{object}: The set of objects to which a token might
    apply. In a grounded plan each Token applies to a specific Object,
    reflecting the intuition that we are using Tokens to describe some
    aspect of an Object (i.e. its state or behavior) in time. However,
    in a partial plan, the commitment to a specific object may not yet
    have been made.

  \item \textit{state}: Tokens can be \texttt{ACTIVE, INACTIVE,
      MERGED}, or \texttt{REJECTED}. The state variable captures the
    token's current state and its reachable states through further
    restriction.  This is required to support CAIP's approach to
    planning where goals and subgoals can be satisfied by either
    adding new intervals (Tokens) to the plan, or by merging with
    existing intervals. Token State lifecycle is examined in detailed
    in a separate section below.

  \end{enumerate}

\item[\textbf{Built-in Object Types}] In a domain model, there may be
  some classes that don't have any time-dependent state or behavior,
  and therefore those classes will not have any Token types associated
  with them, for instance in the Shopping Agent model, the set
  locations and products are static for the agent's purposes so we may
  only want to say what they are but not have any state or behavior
  associated with them.

  However, the most common case in any non-trivial domain model is
  that Objects will have many associated Tokens in order to describe
  their state and behavior throughout a plan. There are a couple of
  Object Types that are so commonly used in domain descriptions, that
  \eu provides a built-in implementation for them:

\begin{enumerate}

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eu's built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{add small
    example?}

\item \textit{Resources}: Metric resources, e.g. the energy of a
  battery or the capacity of a cargo hold, are objects with an
  explicit quantitative state in time and with a circumscribed range
  of changes that can occur to impact that state i.e. produce,
  consume, use, change.  Resources are such a common requirement for
  \eu users that built-in object types (with their corresponding token
  types to denote production, consumption, etc) are provided for
  them. Instances of classes derived from a Resource will induce
  ordering requirements on their Tokens in order to ensure that the
  level of the resource remains within specified limits.  \comment{add
    small example?}
\end{enumerate}

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  Tokens.  Goals can be posted explicitly, for instance, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  compatibilities (described in the CAIP framework above), for
  instance, when the Shopping Agent must get to a Location where a
  Product is available before it is able to purchase it, in this case
  the original goal of purchasing the product will spawn a subgoal of
  being at a specific location.

  Looking at a how a simple Shopping Agent goal would be addresed by
  \eu helps illustrate how Token States support the planning process.

  Let's assume we start with a partial plan that places the Shopping
  agent at home at time 0 and with no Products in its possession, then
  we post a goal for the Shopping Agent to own a Drill at time 10.

  \comment{Add figure showing an AgentLocation timeline with an
    At(Home) token with start=0 and end open, an empty AgentBag
    timeline, and a Own(Drill) token with start=10 floating below both
    timelines}

Notice that the \texttt{Own(Drill)} token that represents the goal is
not immediately placed on the \texttt{AgentBag} timeline since \eu
must determine whether it can achieve that goal and how. When the goal
is posted and before any decisions are made on it, the
\texttt{Own(Drill)} token is created in an INACTIVE state, then \eu
looks at the current partial plan and considers 2 alternatives:

\begin{enumerate}

\item If there are any tokens in the plan that are compatible with the
  new goal, \eu may attempt to satisfy the goal by merging it with any
  of them (if there is more than one it can try them one at a time),
  in this case, the state of the token would change to MERGED.

\item Instead of merging with existing tokens in the plan, \eu may
  decide to insert the new token into the plan, in that case its state
  would change to ACTIVE.

\end{enumerate}

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible with
the inactive token. Two tokens are merge-compatible if they are
instances of the same predicate and that no intersections between
corresponding variables are empty. The effect of merging is
illustrated in Figure 3.

\comment{Add figure showing the effect of merging}

When a token is activated all compatibilities associated with its
token type are evaluated, which may lead to new subgoals being
generated. In our example, there should be a compatibility stating
that the Shopping Agent must be at the same location as the product it
is purchasing. In that case, activating the \texttt{Own(Drill)} token
will cause a new \texttt{At(Drill.location)} to be created in an
INACTIVE state, which in turn will have to be dealt with.

\comment{Add figure showing the effect of activating the token by
  putting it on the AgentBag timeline, and a new At(Drill.location)
  token floating below}

If neither activating nor merging works, a token may become REJECTED,
this is possible only for goals posted by the end user and specified
as optional. Subgoals generated as a result of token activation cannot
be rejected, if a subgoal cannot be satisfied through activation or
merging then it will cause the original activation decision that
spawned it to be retracted (\eu's search mechanisms are explained in
detail in a separate section later on).

As we can see, a token's state can go from INACTIVE to ACTIVE, MERGED
or REJECTED, this lifecycle is illustrated in the figure below.

\comment{Add figure showing possible state transitions}

\item[\textbf{Rules}] In order for a plan to be valid, it must comply
  with all rules and regulations pertinent to the application domain
  in question. Rules govern the internal and external relationships of
  a token. For example, consider a parameterized action
  \texttt{Go(origin,destination)} which describes a Shopping Agent
  traveling from one location ( \textit{origin}) to another (
  \textit{destination}). The parameters are instantiated on a token as
  variables whose domain of values is the set of all locations in a
  given problem. A rule governing an internal relationship among token
  variables might stipulate that a transition must involve a change in
  location. This can be easily expressed as a constraint on the
  definition of a predicate of the form: \textit{origin !=
    destination}. It makes sense to further stipulate that the agent
  must be located at \textit{origin} before travel can happen and the
  agent must be located at \textit{destination} when completed. This
  is an example of a rule governing an external relationship among
  tokens. It specifies a requirement that tokens of the predicate
  \texttt{At} that represent the location state for the Shopping Agent
  precede and succeed tokens of the predicate \texttt{Go} that
  represents traveling.

  \comment{ Add figure showing internal and external relationships
    from rules on action Go}

Figure xx illustrates the entities and relations involved in
specifying such a rule on a \texttt{Go} action. The token on which the
rule applies is referred to (\texttt{Go}) as the master. Each
\texttt{At} token required by the master is referred to as a
slave. All variables are indicated by name and their domains are
expressed as intervals in the case of temporal variables and as
enumerations for the remainder. Application of a rule on a token can
thus cause slave tokens, variables, and constraints to be introduced.

The capability of a domain rule to cause a slave token to be created
is a key vehicle through which planning occurs. Semantically, this
rule imposes a requirement for supporting (slave) tokens to be in the
plan in order for the master to be valid.

\comment {talk about condition/effect annotations?}

\end{description}

We have just covered the main elements of \eu's planning
representation:

\begin{enumerate}
\item \textbf{Domains, Variables and Objects} to represent the
  entities (Objects) and attributes (Variables) needed to represent a
  problem domain

\item \textbf{Tokens} Temporally scoped predicates to represent states
  and actions in time.

\item \textbf{Token States} which are necessary for the implementation
  of planning operations on a partial plan.

\item \textbf{Constraints} to describe relationships among attributes
  of objects and tokens. \eu has a built-in library of constraints
  that covers time, resources and others that are common in real life
  problem domains.

\item \textbf{Domain Rules} to describe internal and external
  relationships on and between tokens respectively. Rules are applied
  to active tokens (master tokens) referred to as masters and
  typically produce inactive tokens (slaves) which are a key vehicle
  through which the planning process evolves. \comment{citation, or
    perhaps remove this statement?:}The rule structure in \eu differs
  from the more restrictive commitment to preconditions and effects in
  classical planning which prohibits durative actions and disembodied
  effects (i.e. effects which may not occur until some temporal
  distance after the end of the action).

\item \textbf{Timelines and Resources} are built in implementations of
  Object types that appear frequently in real life domains.

\end{enumerate}

%\begin{figure*}
%\centering 
%\subfloat[]{\label{fig:plan-evolve1}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-1.pdf}} 
%\subfloat[]{\label{fig:plan-evolve2}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-2.pdf}} 
%\subfloat[]{\label{fig:plan-evolve3}\includegraphics[width=0.3\textwidth]{figs/Plan-evolve-3.pdf}} 
%\caption{\small An illustrative plan synthesis example with concurrent
%  timelines. Each timeline is an instantiation of a subsystem tracked
%  by the planner. Tokens describe the instantiated state of subsystem
%  at a particular time instance and enforce temporal and parametric
%  constraints. Abstract goals are decomoposed into successively less
%  abstract tokens and instantiated in co-temporal timelines using
%  \texttt{Allen Algebra} relations. All tokens represent flexible
%  start/end times. \ref{fig:plan-evolve1} shows an initial state
%  evolving into \ref{fig:plan-evolve2} and \ref{fig:plan-evolve3}.}
%  \label{fig:Plan-evolve}
%  \vskip-5pt
%\end{figure*}

%Fig. \ref{fig:Plan-evolve} shows an illustrative example. We show four
%timelines which track \texttt{Goal, Path, Navigation} and
%\texttt{Command} state over time. Tokens on the goal timeline indicate
%a survey within a bounded volume and a min/max depth envelope. As the
%Volume Survey sub-goals on the \texttt{Path} timeline, the initial
%token on that timeline gets \emph{squeezed} to have dependencies of
%the goal token instantiated. Initially this dependency are the tokens
%\texttt{Go(x1,y1)} and \texttt{Go(xn,yn)} indicative of the area of
%coverage. The first of these in turn generates a sub-goal on the same
%timeline to go to the next intermediate waypoint, which is illustrated
%in Fig. \ref{fig:plan-evolve2}. Subsequent sub-goals (not shown in the
%figure) ultimately generate all the tokens to completion for this
%timeline. Fig. \ref {fig:plan-evolve3} shows a snapshot further along
%in plan generation and shows additional subgoals on the \texttt{Path,
%  Navigation} and \texttt{Command} timelines. Each token can thus
%trace its causality when instantiated in the plan. The domain model
%provides source of these temporal and parametric dependencies in the
%form of rules.

%\begin{minipage}[c]{\textwidth}
%\vspace{+0.5cm}
% \framebox[\textwidth][t]{
%Volume\_Survey(x$_1$,y$_1$,x$_n$,y$_n$,Min\_Depth, Max\_Depth)\\
%$\Rightarrow$ \\
%met\_by Go(x$_1$,y$_1$,Min\_Depth); \\
%starts [50,0] Go(x$_2$,y$_2$,Min\_Depth,Max\_Depth); \\
%\ldots{} \\
%ends [50,0] Go(x$_{n-1}$,y$_{n-1}$, Min\_Depth,Max\_Depth); \\
%meets Go(x$_n$,y$_n$,Min\_Depth, Max\_Depth); 
%\end{minipage}

% \end{enumerate}

% \subsubsection{Partitioned Control}

% \subsubsection{State Estimation}

% Even as automated reasoning approaches have the ability to dynamically
% retarget the vehicle, estimating environmental signals of interest (or
% their proxies) is important to be able to enable opportunistic science
% in the water-column. In \texttt{T-REX} feature recognition revolves
% around a Hidden Markov Model (HMM) \cite{rabiner86} which is encoded
% directly within the unified representational and computational
% framework of a reactor. HMMs are useful since the stochastic nature of
% these models can correlate the type of features we want to detect with
% the sensor observations. Online targeted sensor data is classified by
% apriori generated cluster data which in turn determines the
% probability of having seen the feature of interest. Together with
% posterior probability, the HMM generates a probability of being within
% the target feature if it dominates the distribution. If other sampling
% conditions are satisfied (for instance sampling proximity), a
% constraint is activited in the model which can trigger a water sampler
% and also preserve the memory of sampling to alter a future transect
% resolution.

% The HMM is customized for the feature of interest and is built offline
% using clustering techniques from large data sets; we use Kohonen's
% Self Organizing Maps \cite{kohonen}. We use a semi-supervised learning
% method \cite{zhu05} to extract an environmental model that make uses
% of both labeled and unlabeled data for training. Online,

% The sensor data for classification depends on the
% feature of interest; for detecting INLs we use backscattering data
% from a HydroScat and for tracking riverine plumes to track fertilizer
% runoff, we use an ISUS Nitrate sensor.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ieee-ram09"
%%% End: