### 2012-02-16 europa.tex
citations added to europa text

\subsection{\eu Architecture}
\label{sec:europa:arch}

\eu is implemented as a library that encapsulates all the elements described above and makes them easily accessible through modeling and API layers.
A user will typically use the modeling facilities to inject a problem domain description, and the API to extract and manipulate plans, and to obtain information about plan flaws and constraint violations at any stage of the planning process.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figs/EUROPA-Architecture.jpg}
\caption{\small The \eu Architecture}
\label{fig:europa-architecture}
\vskip+0.1cm
\end{figure}

Fig. \ref{fig:europa-architecture} shows the main architectural
components in \eu and their relationships.

The \textbf{Constraint Reasoning Engine} (CRE) manages domain types, variables and constraints that define
relationships among them. It also provides an efficient arc
consistency mechanism \cite{mackworth77}. The CRE is designed so that
specialized reasoning algorithms for specific constraints can be easily
and efficiently plugged in.

The \textbf{Plan Database} (PDB) manages object and token types. It also maintains the state of partial plans in terms of object, variable and token instances. This constitutes the data foundation on which end user applications and automated planners are built.

The \textbf{Rules Engine} (RE) modularly adds the capability to support domain rules that describe dependencies between tokens,

The \textbf{Solver Module} provides the framework for flaw detection and resolution that is needed to implement search algorithms for planning.

The \textbf {Language Modules}: Describing problem domains and problem instances in terms of objects, variables and domain rules in a programmatic way is cumbersome to the extent of being impractical, that motivated the creation of modeling languages that support those definitions in a much more natural and compact way. NDDL is the main modeling language for \eu and will be described in detail in the Modeling section below. ANML is a language under development that provides a more intuitive syntax \comment{provide ANML citation}.

The \textbf{API Layer}  provides access to services from all the different modules in a programmatic way so that \eu can be easily embedded into specialized applications.

In an effort to mimimize the amount of work that a user needs to perform to solve specific problems, \eu  also provides extension modules that have been found to be useful in many domains: 
\begin{enumerate}
	\item \textit{Temporal Network}: Extensions to the CRE module to reason about temporal constraints 
	\item \textit{Resources}: Extensions to the CRE, PDB and Solver modules to provide out-of-the-box reasoning and search for metric resources.
	\item \textit{Chronological Backtracking Solver}: A general purpose built-in solver to provide the ability to plan out-of-the-box, problems that require challenging search may require specialized solvers to be built as explained in the Search section below.	
\end{enumerate}

\subsection{Modeling}
\label{sec:europa:modeling}

\eu can ingest descriptions of models, plans and goals and thereby can
be applied to different domains and problems, simply by providing
different models and goals. Consequently the expressiveness of the
language for models, goals and plans is of great importance. For \eu,
NDDL (New Domain Description Language [pronounced ``noodle'']
allows the user to specify the different components of a
problem domain in a concise way. The main characteristics that make NDDL
a powerful tool for describing problem domains are:

\begin{enumerate}

\item Object Oriented: NDDL supports classes and
  inheritance in similar fashion to popular Object-Oriented languages
  like Java and C++. It also supports polymorphism for some of the
  planning components. Using object classes and instances is a
  time-tested approach to naturally describe problem domains \comment{OO
    citation}.

\item It offers declarative constructs to define constraints
  and causality dependencies between actions and effects.

\item It offers procedural constructs to populate \eu's plan database
  with specific problem instances expressed in terms of objects,
  variables, constraints, facts and goals.

\end{enumerate}

To understand how the most important elements of NDDL work, let's take a
look at how the Shopping Agent problem could be modeled:

\begin{verbatim}

// Locations (Home, SuperMarket, etc.)
class Location {
  string name;
  Location(string _name){
    name = _name;
  }
}

// Products (Milk, Banana, etc.) and
class Product {
  string name;
  Product(string _name) {
    name = _name;
  }
}

// ProductLocations (Banana can be found at SuperMarket, for example)
class ProductLocation {
  Location location;
  Product product;

  ProductLocation(Location _location, Product _product){
    location = _location;
    product = _product;
  }
}

// Use built-in Timeline functionality to enforce that an agent:
// a) Can't be at more than one place at a time.
// b) Can't Go more than one place at a time.
// c) Can't Go somewhere and be At somewhere at the same time.
class AgentLocation extends Timeline{
  predicate At {
    Location loc;
  }

  predicate Going {
    Location from;
    Location to;
  }
}

// In addition to having a location timeline, the agent can buy and own things.  
// Note that the actions and the location predicates can be concurrent 
// so they can't be on the same Timeline
class Agent {
  AgentLocation location;

  Agent() {
    location = new AgentLocation();
  }

  action Buy {
    Product product;
  }

  action Go {
    Location from;
    Location to;
  }

  predicate Own {
  	Product product;
  }
}
\end{verbatim}

In the model above we can see how the main elements or a problem domain can be mapped to \eu's representation:

\begin{enumerate}
	\item \textbf{Object Types}: NDDL supports an object-oriented approach, where common data and behavior can be abstracted as a class. In this case, we have the \textit{Product} class represent the products that the Shopping Agent needs, \textit{Location} represents the places where products can be found, \textit{ProductLocation} represents the association arising when a product can be purchased at a location. The Shopping Agent itself is represented by an \textit{Agent} class, which relies on the \textit{AgentLocation} class to keep track of where the agent is at any point in time. NDDL supports a single rooted class hierarchy so that state and behavior can be extended through inheritance. We took advantage of that by making \textit{AgentLocation} inherit from \eu's built-in \textit{Timeline} class which doesn't allow predicates to overlap in time for the same object instance, in this way we can ensure that and agent can only be at one place at a time. 
	\item \textbf{Variable Types and Instances} : All object attributes are represented as variables in this model, for instance, it uses integer variables to represent the built-in temporal parameters for each token (activity duration, start and end times), string variables to represent  \textit{Location} and  \textit{Product} names, and object reference variables to represent the structural relationships between different objects (for instance, each Agent has variable to keep track of its  \textit{AgentLocation}).
	\item \textbf{Predicate Types}: Some classes like \textit{Product}, \textit{Location} and \textit{ProductLocation} are completely static and so don't have any predicate or action types associated with them. The \textit{AgentLocation} class is purely used to keep track of state and so it has \texttt{At} and \texttt{Going} predicate types associated with it. The \textit{Agent} class keeps track of the products it purchases through the \texttt{Own} predicate, having a separate \textit{AgentBag} class with its own predicates would have been an equally valid modeling choice.
	\item \textbf{Action Types}: The \textit{Agent} class is the only one that has behavior associated with it, consequently, it defines action types to represent when the agent is taking action to \texttt{Go} to a \textit{Location}, or to \texttt{Buy} a \textit{Product}. We'll see how the actual behavior for those actions is specified below.
\end{enumerate}

\begin{verbatim}

// Define the rules for our actions:
Agent::Go {
  met_by(condition object.location.At origin);
  eq(from, origin.loc);

  equals(effect object.location.Going going);
  eq(going.from, from);
  eq(going.to, to);

  meets(effect object.location.At destination);
  eq(to, destination.loc);
}

Agent::Buy {
  // A Buy takes 10 time units
  eq(10, duration);

  // initialized to all locations
  ProductLocation possibleStores;

  // limit possibleStores variables to ones that provide what we need to buy
  eq(product, possibleStores.product);

  // We must be At a location during our Buy, and that location must have the
  // product we want available:
  contained_by(condition object.location.At currLocation);
  eq(currLocation.loc, possibleStores.location);

  starts(effect Own purchase);
  eq(purchase.product,product);
}
\end{verbatim}

Domain rules can be defined in NDDL in the context of a Predicate or an Action. In the example above we can see the main elements of a rule definition:

\begin{enumerate}
  \item \textbf{Variables}: A Token Type (predicate or action) is part of a class, therefore class attributes (see references to \texttt{object.location} above), Token Type parameters (see references to \texttt{[from, to]}  in \texttt{Agent::Go}, or to \texttt{[duration, product]} in \texttt{Agent::Buy}  and locally declared variables (see \texttt{possibleStores} in \texttt{Agent::Buy}) can all be used in the definition of a Rule. 
  \item \textbf{Constraints}: restrictions on the values that variables can take inside a Rule are expressed through constraints, in the example above only the \texttt{eq} constraint is used, but in general, any constraint available in \eu's library can be used.
 \item \textbf{Subgoals}: In \texttt{Agent::Go}, subgoals are used to specify that the Shopping Agent's location must transition from origin to destination in a temporally consistent way. The \texttt{[meets, equals, metby]} operators are used to specify subgoals (Allen operators are explained in more detail below), along with the temporal relationship each token representing a subgoal has with the the token representing the \texttt{Agent::Go} action. Also, \texttt{[condition, effect]} annotations can be used when defining a subgoal inside an action, \eu doesn't immediately associate any semantics with them, but they can be extremely useful for search modules when creating plans. 
\end{enumerate}

Now, having that we have a specification for the Shopping Agent domain, here is how a particular problem instance could be specified:

\begin{verbatim}

// Allocate instances
Location Home = new Location("Home");
Location SuperMarket = new Location("SuperMarket");
Location HardwareStore = new Location("HardwareStore");

Product Banana = new Product("Banana");
Product Milk = new Product("Milk");
Product Drill = new Product("Drill");

ProductLocation bananaLocation = new ProductLocation(SuperMarket, Banana);
ProductLocation milkLocation = new ProductLocation(SuperMarket, Milk);
ProductLocation drillLocation = new ProductLocation(HardwareStore, Drill);

Agent agent = new Agent();

// Indicate that the database is closed - no new objects can be created
// this allows EUROPA to perform more efficient inference
close();

// We start the day at Home:
fact(agent.location.At atHomeForBreakfast);
atHomeForBreakfast.loc.specify(Home);

// Goals for all of the agent's needs, buy if needed
goal(agent.Own gotMilk);
gotMilk.product.specify(Milk);
goal(agent.Own gotBanana);
gotBanana.product.specify(Banana);
goal(agent.Own gotDrill);
gotDrill.product.specify(Drill);

// Make sure agent is home for dinner
goal(agent.location.At atHomeForDinner);
atHomeForDinner.loc.specify(Home);

// Agent has all day to satisfy goals:
gotMilk after atHomeForBreakfast;
gotMilk before atHomeForDinner;

gotBanana after atHomeForBreakfast;
gotBanana before atHomeForDinner;

gotDrill after atHomeForBreakfast;
gotDrill before atHomeForDinner;

// Force things to happen within our planning horizon:
int horizonStart = 0;
int horizonEnd = 100;
leq(horizonStart, atHomeForBreakfast.start);
leq(atHomeForDinner.end, horizonEnd);

\label{eu:model}
%\caption{Specifying a problem instance for the Shopping Agent in NDDL}
\end{verbatim}

A problem instance consists mainly of 
\begin{enumerate}
  \item \textbf{Variables} and \textbf{Objects}: In this case we defined temporal variables that are used to specify the planning horizon (see \texttt{horizonStart} and \texttt{horizonEnd}), and objects are used to specify the locations, products, product-loction associations and the Shopping Agent that we want to plan for.
  \item \textbf{Facts}: These are used to specify the current state of the world and don't need any support. Facts are represented as tokens in \eu's Plan Database. In this case we use facts to specify that the Shipping Agent is initially at home.
  \item \textbf{Goals}:  Goals are represented as tokens in \eu's Plan Database, differently from Facts, Goals trigger inference in \eu which determines whether flaws (unresolved subgoals, overloaded resources, etc., flaws will be covered in detail in the search section below) need to be resolved, if that is the case, search modules can be invoked which will look for a plan where all the flaws are resolved.
  \item \textbf{Constraints}: restrictions on and between actions and goals are specified through constraints, in this case we use temporal constraints to specify when the facts hold and when the goals should be accomplished. As always, all the constraints in \eu's library (built-in, or user defined), can be used to specify these restrictions.
\end{enumerate}

\subsection{Inference}
\label{sec:europa:inference}

\comment{explain how inference is used to detect perform bounds
  propagation, domain reduction and to detect constraint violations.} 
- talk about AC3 implementation and extension mechanisms
- talk about europa's constraint library

The Temporal Network: The notion of time is central to temporal
planning. \eu uses variables to explicitly represent timepoints for plan
activities and states. Constraints among timepoints provide a natural
way to express domain axioms. For example, in order to state that
activity A must occur before activity B we can say that the end
timepoint of A is $leq$ the start timepoint of B.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} relations shown above.}
\label{fig:allen-algebra}
\vskip-0.3cm
\end{figure}

\cite{dechter91} proposed that constraints among timepoints can be
grouped together to form a Simple Temporal Network (STN). Such a network
can be transformed into a Distance Graph (DG) where the outward arc from
a node represents the maximum distance from the source node to the
target node. They also showed that shortest path algorithms could be
used to propagate values in the network and discover a negative cycle. A
negative cycle is a path from a node to itself that has a path length
less than 0. If such a cycle exists, the network is inconsistent. It was
further shown that a single-source shortest path algorithm was
sufficient to detect a negative cycle and provide sufficient propagation
to yield a backtrack-free search. Thus we have an efficient and complete
algorithm for propagating an STN. These results build on the already
established notion of a CSP and are naturally incorporated into the
general representation and propagation scheme used in \eu.
The diagram below illustrates a simple STN with just 2
variables and a single constraint. It also shows the resulting DG.

\begin{figure}[!htb]
  \centering
  \subfloat[\small A Simple Temporal Network (STN)]{\label{fig:stn1}\includegraphics[width=0.4\columnwidth]{figs/stn-graph.pdf}}
  \subfloat[\small Resulting distance graph from \ref{fig:stn1}.]{\label{fig:stn2}\includegraphics[width=0.4\columnwidth]{figs/distance-graph.pdf}}
    \caption{\small An illustration of a STN and its corresponding
      distance graph.}
  \label{fig:stn}
\end{figure}

\comment{ add figure. Use Europa wiki figure; or use conor figure from slides}

Resources:

\begin{enumerate}
  \item \textbf{Reservoir} resources can have both 'consume' and 'produce' tokens. Note
that these tokens do NOT have the usual start/end/duration token
variables, since occur at a single instant in time. Instead, they have a
single variable, 'time' used to represent that instant.

  \item \textbf{Reusable} resources have 'uses' tokens that use a quantity for the
duration of the token. These tokens do have the usual start/end/duration
variables.

  \item \textbf {Unary} resources are reusable resources with unit quantity and have 'use'
tokens (again, with the usual start/end/duration variables).
\end{enumerate}

\comment{talk about resource envelope computation by using MaxFlow formulation.}

\subsection{Search}
\label{sec:europa:search}

Now we have all the elements in place so that an automated problem
solver can be created. Let's recap what those elements are:

\begin{enumerate}

\item A domain model that describes the variable, object, predicate and
  action types that are relevant for the problem.

\item A problem instance (also called initial state) that consists of:
	\item Variable and Object instances that exist for the entire planning horizon.
	\item Temporally scoped predicate and action instances.
	\item Temporally scoped goals.

\end{enumerate}

All this information is kept in \eus Plan Database so that inference and
search mechanisms can be used to look for a problem solution.

Given that temporal intervals may extend indefinitely in both positive
and negative directions, for problems with a temporal dimension it is
also common to specify a Planning Horizon when invoking a solver. The
Planning Horizon is the temporal interval that a problem solver has to
worry about, any decisions that fall completely outside of the Planning
Horizon can be ignored. This is very useful for recurrent activities,
consider for instance a model where a medical checkup has to occur every
3 months, the user can model a recurring activity and then use the
Planning Horizon to ensure that a finite number of medical checkup
activities are generated as part of a plan. This is also useful to
ignore activities that may happen too far in the future or in the past
to be relevant for a particular plan, this way the user can ensure that
the solver is making as few commitments as necessary.

\eu provides a built-in solver that performs Plan Space Planning
\cite{ghallab04} in this approach, the initial state is considered a
partial plan that needs to be refined toward a solution plan that
achieves the goals. The operations to refine the partial plan PP at any
time are:

\begin{enumerate}
	\item Find the flaws of PP, that is the conditions that prevent it from being a solution plan (flaw types are explained below).
	\item Select one such flaw
	\item Select a resolver for the flaw
	\item Refine PP by applying the resolver
	\item If an inconsistency is found, try another resolver.
	\item If all possible resolvers for a particular flaw fail, return failure, otherwise continue until resolving all the flaws.
\end{enumerate}

This Plan Space Planning algorithm is translated into \eu's
representation as follows:

The initial partial plan is the state of \eus Plan Database after the
initial state has been instantiated, this results in a set of variable,
object and token instances. Inference then takes place to detect flaws
in the partial plan, there are 3 kinds of flaws that can be detected:

\begin{enumerate}

\item Unbound Variable: a variable in the partial plan whose domain is
  not a singleton. Unbound Variables are resolved by specification of a
  value from the domain of the variable.

\item Open Condition: an open condition is an inactive token. Inactive
  tokens can be generated by explicitly posted goals, or when a token is
  activated, rules may fire that create inactive slave tokens
  \comment{point to more detailed decription above in Rules Engine
    section}.
  Open
  Conditions can be resolved in three ways:

  \begin{enumerate}

  \item Merging: A free activity is merged with a matching activity
    already in the plan. An activity a is said to match an activity a?
    if a and a? unify and the temporal constraints involving a are
    satisÞed by a?. Thus, a and a? can be considered the same activity
    and we do not need to introduce a in the plan. Consequently, the
    compatibilities associated with a are not Þred, because they have
    been already triggered when a? was introduced in the plan.

  \item Activation: We introduce a new activity a in the current plan
    associating it with the proper timeline, but without choosing a
    speciÞc time slot for it.The compatibilities associated with a are
    applied and the subgoal activities resulting from those
    compatibilities are introduced as free activities. This results in
    both an ordering flaw, corresponding to the just activated activity,
    and a number of open condition flaws, corresponding to the new
    subgoal activities.

  \item Rejection: Some goals may be optional, if that is the case for a
    particular order condition, its corresponding flaw can be resolved
    by discarding the inactive token.

\end{enumerate}

\item Threat: once a token has been placed in the partial plan it may
  impact other tokens indirectly through possible overlapping
  requirements on objects. Recall for example that a token may belong to
  objects (e.g. Timelines) which require a total order over their
  tokens. If any 2 tokens could possibly overlap (though not
  necessarily), then they pose a threat to each other in terms of
  achieving an extension of the current partial plan which is complete
  and consistent. Similarly, threats may arise where tokens share a
  common resource and their current state might yield extensions of the
  current partial plan which are inconsistent. Threats are resolved by
  imposing ordering constraints among tokens.

\end{enumerate}

Open Conditions and Threats allow flaw detection and resolution at a
higher-level of abstraction (i.e. in terms of objects and tokens) than
that of simply binding variables as is common in Constraint Satisfaction
Problems. This is advantageous when one applies heuristics for ordering
choices since it provides a richer context in which to make decisions.
Furthermore, it aids in reducing the amount of work done by a solver so
that only the necessary refinements are made, otherwise leaving the
partial plan with maximum flexibility. For example, one can omit unbound
variables which are time-points of tokens since threats will force a
solver to impose restrictions on these variables based on the semantics
of the objects to which their tokens apply. Thus the planning process
may yield partially-ordered plans for which all possible extensions are
provably valid.

The Plan Space Planning algorithm described above can be implemented in
many different ways depending on the approach chosen for flaw and
resolver selection and for backtracking. \eus built-in solver implements
a chronological backtracking algorithm that is summarized in Algorithm
\ref{alg:europa:solve}.

\begin{algorithm}[H]
\KwIn{PartialPlan $plan$}
\KwOut{\texttt{true} if a plan solution was found, \texttt{false} otherwise}
\Begin{
  \lIf{isInconsistent(plan)}{\Return \texttt{false}\;}\label{li:propagate}
  \BlankLine
  Flaw $flaw \leftarrow chooseFlaw(plan)$ \; \label{li:selectflaw}
  \lIf{ $\emptyset = flaw$ }{\Return \texttt{true} \;} \label{li:noflaw}
  \BlankLine
  DecisionPoint $decision \leftarrow makeDecisionPoint(flaw, plan)$ \; \label{li:decisionpt}
  \While{ $decision.hasNext()$ }{ \label{li:decisionloop}
    PartialPlan $pp \leftarrow decision.executeNext()$ \; \label{li:decide}
    \lIf{ $solve(pp)$ }{ \Return \texttt{true} \; } \label{li:recurse}
    \lElse{ $decision.undo()$ \; } \label{li:backtrack}
   }
   \Return \texttt{false} \; \label{li:noplan}
 }
\caption{$\mathrm{bool} ~ solve(plan)$}
\label{alg:europa:solve}
\end{algorithm}

The algorithm takes as input a partial plan p and returns true if a
complete and consistent refinement of p could be found (or if p is
initially complete and consistent), and false otherwise.

\begin{enumerate}
	\item Line \ref{li:propagate}: Propagate the constraints to test for inconsistency. If found to be inconsistent, then we can return false since this is a dead-end i.e. no refinements to p can yield a consistent plan.
	\item Line \ref{li:selectflaw}: Choose a flaw from the set of available flaws.
	\item Line \ref{li:noflaw}: If there are no flaws, then p is complete and we can terminate with success.
	\item Line \ref{li:decisionpt}: Otherwise we formulate a decision point which is a branch in the search space. Each choice is a particular refinement operation and the DecisionPoint collects all possible refinement operations for the given flaw.
	\item Line \ref{li:decisionloop}: We keep trying until the chosen flaw is resolved, or until we run out of resolvers to try.
	\item Line \ref{li:decide}: A new partial plan is obtained by application of a refinement operator. Note that the ordering over refinement operators to select is a non-deterministic step.
	\item Line \ref{li:recurse}: Recursive call to solve the new planning problem. If successful, then we are done.
	\item Line \ref{li:backtrack}: Otherwise, retract the last refinement operation and move on to try the next one.
	\item Line \ref{li:noplan}:  If we arrive here, then we have exhausted all options to resolve the flaw, including the case where no options were available initially. Thus the problem cannot be solved.
\end{enumerate}

This algorithm provides for a sound and complete search, assuming that
no flaws or available refinement operators are pruned unnecessarily.
Dead-ends in the search are discovered through constraint propagation.
Constraint propagation is a vehicle for evaluating the consistency of a
partial plan and also for filtering infeasible values from consideration
prior to commitment, allowing in some cases a strong look-ahead
capability which is essential for tractable search. Consistency testing
is initiated by the isConsistent procedure. The algorithm permits a
heuristically controlled search by applying orderings for chooseFlaw and
makeDecisionPoint. It results in a chronologically-backtracking search.

\comment{expand a little more on how heuristics can be implemented by
  applying ordering/filtering for chooseFlaw and makeDecisionPoint}
examples:
- deal with earliest flaws first
- deal with planning (open condition flaws) before scheduling (threats) ones

It should be emphasized that while this algorithm is commonly employed,
it is only one of many that could be implemented. \comment{give examples
  of local search solvers for scheduling and CSP?}

N-Queens solver:
\begin{verbatim}
    public void solve(int maxIter)
    {
    	init();

    	for (int i=0;psengine_.getViolation() > 0 && i < maxIter;i++) {
    	    PSVariable queenToMove = getQueenWithMaxViolation();
    		int curPos = queenToMove.getSingletonValue().asInt();

    	    boolean moved = false;
            SortedSet<Move> moves = getMoves(queenToMove,curPos);             
            for (Move m : moves) {
            	moved = makeMove(queenToMove,curPos,m,false);
            	if (moved)
            	    break;
            }

            if (!moved) 
            	makeMove(queenToMove,curPos,moves.first(),true);

            checkSolution(); // See if we have a new best solution            
            notifyIterationCompleted(curIteration_++);

            if (curIteration_-bestIter_ > 50) 
            	restart();
    	}
    }
\end{verbatim}

RCPSP solver:
\begin{verbatim}
    public void solve(PSEngine psengine,
    		          long timeout, // in msecs
    		          int bound)
    {
       init(psengine,timeout,bound);

       for (int i=0; true ; i++) {
          flatten();
          updateSolution(i);
          updateCriticalPrecedences();

          if ((nbStable_ > maxStable_) || (bestMakespan_ <= makespanBound_))
             break;

          if (timer_.getElapsed() > timeout) {
              timedOut_ = true;
              break;
          }

          relax();
          curIteration_++;
       }

       restoreBestSolution();
    }

\end{verbatim}

\comment{explain how other solvers can be built, maybe use NQueens or
  RCPSP example to illustrate. Or maybe build a little specialized
  solver for shopping or rover examples.}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "setobook"
%%% End: