### 2012-02-24 future.tex
final updates

\section{Future Work}
\label{sec:future}

Our experience of using deliberative techniques in a highly
inter-disciplinary environment for targeted marine science
applications has shown the need and applicability of such novel
algorithms and methods. In addition to augmenting traditional AUV
surveys more advanced Lagrangian observations, mixed-initiative
methods, our colleagues at MBARI have understood the longer term
implications of our work. Our current and consequently future efforts
are therefore grounded in important science problems in an operational
oceanographic setting.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{figs/autonomy-topics.pdf}
  \caption{\small Overlapping topics in our research in deliberation
    and autonomy with sampling the dominant theme.}
  \label{fig:topics}
\end{figure}

The core of our problem domain for \can has been towards Sampling. In
large part given the spatial and temporal scales of the
bio-geochemical processes we are interested in in this project, it is
imperative that a key direction is in formulating a multi-robot
adaptive sampling problem. In this context while superficially \rx can
be considered as a multi-agent system and could therefore imply a
direct extension, sporadic communication with high-latencies in
surface communication, not to mention low throughput acoustic
communications imply that the extension of \rx for distributed
applications is non-trivial. Further, in order to efficiently execute
coordinated observation of ocean processes with multiple autonomous
sensing platforms, significant methodological as well as technical
challenges must be addressed. We are working towards a systematic
approach of using \rx as a backend planner for \can field
experiments. The expectation is that \od users will plan initially
individual vehicle surveys which are variants of standard
straight-line transects. Lessons learned with human planners in the
loop will be used towards planning larger surveys, also with simple
fixed transects but using vehicle capabilities in ontologies similar
to \cite{patron09}. Apportioning the planning task between shore-side
components with substantial contextual information and compute power
with robots with in-situ reasoning capabilities a la \rx is a
challenge and an open research problem. Fig. \ref{fig:topics} shows
what we believe is the scope and influence of the different topic
areas we are focusing on.

Partitioned inference and decision-making within \rx also lends itself
to graceful system degradation, making the overall approach effective
towards increased robustness. There are two outstanding diagnosis and
failure recovery challenges that we need to address; the first has to
do with methodology related to model design of the instrument payload
for a highly modular and configurable vehicle; this is local to a
reactor. Doing so would enable (as a first step) using systematic
Fault Diagnosis, Isolation and Recovery (FDIR) algorithms as
demonstrated in the \texttt{RAX} experiment
\cite{williams96,mus98,williams97} and more recently on an AUV by
Dearden and others \cite{wang09,ernits10,dearden11}. The second
challenge has to implementing a policy within the \rx framework that
explicitly deals with recovery since controller failure can be
distributed across reactors. Since reactors are hierarchical in their
relationship to one another, their well defined dependencies allow
them to be removed from an agent systematically. Reactors more
abstract in scope can be removed before those which are less abstract
(and closer to the hardware). In the event a Navigation reactor, for
example, has a fault, it can be removed safely with lower level
functionalities (such as those in the executive). When the Navigator
is removed, it will recursively remove more abstract reactors (such as
a Mission Manager) that depend on it. Such a safety feature then
allows the system designers to ensure that early on they can invest
more in ensuring that reactors lower in the hierarchy are robust and
potentially formally verified. It also means that there is a safe and
sure way to take control of the vehicle at any level within the
reactors hierarchy; as long as low level reactors are still active one
will always been able to send goals/commands to them and execute what
are traditionally known as runout sequences (or clean up scripts).

Feature recognition for Sampling is a general research thrust; however
it has specific relevance in the context of \rx and AUV control. As
noted in Section \ref{sec:results} we have already demonstrated
Machine Learning techniques for recognition of select features in the
coastal ocean. A principal challenge that remains is that of obtaining
expert labeled data for any supervised or semi-supervised learning
methods.  While supervised techniques for learning classifiers such as
decision trees (\cite{Quinlan93-dtrees}), instance-based
classification (\cite{Aha-ibl-ml91}), Bayesian classifiers
(\cite{Jensen2001-BNetworks}) and artificial neural networks
(\cite{ANNsurvey-2000}) are popular, given the vast amount of data
available and poor understanding of correlation between physical and
biogeochemical variables in the coastal ocean, our work on INLs
\cite{mcgann08d,ryan10} and reliance on a oceanographic expert has
shown that scaling to different problems remains problematic. This is
a prime motivation to move towards more semi-supervised methods such
as \cite{kumar11,sergio12}. With the advent of the \od however,
another direction our research is taking us is in \texttt{Recommender
  Systems} \cite{Adomavicius05} as noted in Section
\ref{sec:results}. We are working to design and deploy software
infrastructure that will learn from user tagged data. To do so, we
will use tagging techniques similar to what users of commercial image
sharing sites like Flickr and Picassa do, to learn the context of the
data (including remote sensing). In addition, the \od will integrate
statistical Machine Learning techniques with incoming asynchronous
data stream obtained by sensors and platforms. Identification models
will be based on existing training sets for INLs and plumes; models
for other features of scientific interests driven by \can requirements
will be built in addition. Existing AUV archives for instance provide
sufficient data for a learning base for a range of features of
interest to \can; in addition we will build software that will learn
to incorporate incoming data streams to provide an existence proof of
a feature of interest.

A number of issues remain to be explored improving \rx and \eu
integration particularly in the context of engineering models for
execution.

Dispatchability of the {\em external} temporally flexible partial
plans continues to be an open research problem. Substantial progress
has been made \cite{vidal97,mus98a,tsam98,morris00} by analyzing the
plan structure of static offline temporal plans. However in the
context of embedded plan execution, balancing model design while
dealing with execution uncertainty requires further research. For
instance it should be possible to distinguish part of the plan that
contributes to a reactor goal, which often needs to be dispatched as
early as possible with other tokens describing world evolution with a
preference to dispatch as late as possible. Making this distinction
would improve the quality of agent behavior while avoiding extra model
complexity.

More abstractly at the architectural level, the ability to dynamically
manipulate the reactor graph would be of importance while creating new
reactors or modifying timelines. Doing so would provide more robust
failure recovery mechanisms especially after a reactor has failed to
synchronize. For example, an alternate reactor could take ownership of
timelines left invalid by the destruction of another failed reactor
ensuring a transparent implementation of component redundancy. While
the current design broadly provides for such a capability, the \rx API
has yet to implement it while ensuring other design elements within
\rx and \eu are not compromised.

Finally and importantly, integration of reasoning about resources
within \rx and tying it to existing \eu capabilities needs to be
demonstrated. Specifically \rx needs to be able to update resource
levels as incoming observations.  Doing so would ensure plan
adaptation occurs dynamically when planned and observed resource
values diverge -- for instance if the vehicle looses a bank of
batteries during the course of a mission.  Enhanced modeling
capabilities and augmented domain-independent search within \eu would
make it easier for \rx users to create such domain models and to
perform more proactive deliberation.

Along with all these potential extensions, we are also very cognizant
that the learning curve associated with \rx and \eu for a general user
in the marine robotics community needs to be tackled. At present, the
complex system that underlies \rx and its underlying framework, even
if well documented, can be intimidating for a naive first time user.

% multi-vehicle control
% sampling
% diagnosis and recovery
% feature recognition
% scalable architecture for missions

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "setobook"
%%% End: