### 2012-04-20 setobook.tex
updates

% \documentclass[a4paper, 12pt]{article}
\documentclass[graybox,14pt]{svmult}

\input{preamble}
\setcounter{tocdepth}{3}

% \titlerunning{Towards Deliberative Control of Autonomous Robots}
% \authorrunning{K.Rajan et. al.}

\title*{\textbf{\sc Towards Deliberative Control in Marine Robotics}}
\author{\normalsize Kanna~Rajan\inst{1} \and Fr\'ed\'eric Py\inst{1}
  \and Javier Barreiro\inst{2}}
\institute{Monterey Bay Aquarium Research Institute, California
  \emph{\{Kanna.Rajan, fpy\}@mbari.org}
  \and
  NASA Ames Research Center, Moffett Field, California
  \emph{javier.barreiro@nasa.gov}}

\begin{document}

\tableofcontents{}

\maketitle

\abstract{We describe a general purpose Artificial Intelligence based
  control architecture that incorporates in-situ decision making for
  autonomous underwater vehicles (AUVs). The Teleo-Reactive EXecutive
  (\rx) framework deliberates about future states, plans for actions
  and executes generated activities while monitoring plans for
  anomalous conditions. Plans are no longer scripted a priori but
  synthesized onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation. \rx is in routine operational use at MBARI, providing
  scientists a new tool to sample and observe the dynamic coastal
  ocean.}

\input{intro.tex}

\input{related.tex}

\input{europa-basics.tex}

\input{concepts.tex}

\input{europa.tex}

\input{arch.tex}

\input{results.tex}

\input{future.tex}

\input{conclude.tex}

\section{Acknowledgements}

Rajan and Py are funded by a block grant from the David and Lucile
Packard Foundation to MBARI. They thank their collaborators John Ryan,
Julio Harvey, Jim Bellingham and Robert Vrijenhoek at MBARI. In
addition they thank colleagues Martha del Alto, David Korsmeyer and
Steven Zornetzer at NASA Ames who made \eu available as open
source. The authors thank other members of the \eu team over the
years, including Ari J\'onsson, Jeremy Frank, Conor McGann, Paul
Morris, Michael Iatauro, Matthew Boyce and Tristan Smith.  Minh Bin Do
(NASA Ames), William Yew Teck Tan (NUS, Singapore) and Rishi Graham
(MBARI) provided comments on early drafts; we are grateful for their
help in making this manuscript more accessible. Finally, the authors
are indebted to Nicola Muscettola for his ideas and early efforts in
Constraint-based Temporal Planning at NASA Ames which forms the basis
of much of this work.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}