### 2012-05-04 Review-response.tex
updates

\documentclass[11pt,twoside]{article}
% \usepackage{subfigure}
% \usepackage{latexsym}
% \usepackage{graphicx}

% \usepackage{floatflt}
% \usepackage{float}

\input{preamble}

\renewcommand{\baselinestretch}{0.973}

\begin{document}
\thispagestyle{empty}
\pagestyle{empty}
\setlength{\parindent}{0pt} 
\setlength{\parskip}{2ex}

\small{Kanna Rajan}\\
\small{Principal Researcher for Autonomy}\\
\small{Monterey Bay Aquarium Research Institute}\\
\small{Moss Landing, California}\\
\small{http://www.mbari.org/staff/kanna/}\\
\small{http://www.mbari.org/autonomy/index.htm}\\
\small{(831) 775-1833}\\

Reviewer and Editor for the manuscript \textsf{\underline{Towards Deliberative Control in Marine Robotics}}\\

Dear Mae,

First and foremost, we would like to thank the reviewer of our
manuscript above for the extensive and valuable comments provided. And
we thank you for your efforts to pull this all together.

We are heartened by the reviewers comments that the manuscript
``follows the intent and spirit'' of the book's theme. And we are
delighted in her/his belief that the manuscript should be accepted for
publication.

We have \emph{carefully} gone over all the comments provided. The
attached revised manuscript reflects those questions/comments from the
review and are highlighted in \kcomment{red} . In a number of
instances, we have taken the liberty to tighten the text for the
English and in one instance added some suplemental material which is
substantial, to explain how the \eu planner components work. We have
also added a couple of figures to better illustrate the concepts per
the reviewer's suggestions. The changes in sum are \underline{not}
substantial given the feedback. However the page count has, with the
changes, gone up a little. And as you would likely see, we've worked
hard to address all the comments while keeping the material
accessible.

Thank you once again for all your efforts on our collective behalf for
this book. We look forward to working with you to bring the book out
for educators at the earliest.

\vspace{+2cm}
\begin{tabbing}
\hspace{+10cm}\=\small{Kanna Rajan}\\
\>\small{Corresponding Author}\\
\end{tabbing}

\end{document}
\end{input}