### 2012-06-05 final-version/conclude.tex

\section{Conclusion}
\label{sec:conclusion}

% Final parting words summarizing the contribution of deliberation towards
% vehicle autonomy.

The Ocean Sciences are at a cusp, straddling traditional ship-based
exploration with newer observatory-based methods. In the United States
the NSF funded Ocean Observatories Initiative \cite{ooi} and regional
bodies such as \texttt{CeNCOOS} \cite{cencoos} are investing in new
observational methods and technologies which \kcomment{aim} to aid
oceanographers by producing substantially higher-resolution data
products. These will use moorings, cabled observatories and glider
fleets generating near real-time data from one or more locations as a
continuous stream. The tack chosen by \can has been to complement such
methods with the aim of developing methods for a \emph{portable
  mobile} observatory which is dependent on autonomous platforms. In
this context what we advocate in this chapter is an initial step in
making robotic platforms more adaptive, scaling them towards fleets of
vehicles (underwater \kcomment{and/}or on the surface) and most
importantly using them to solve the larger challenge of sampling
\kcomment{dynamic} spatio-temporal fields.

While our trajectory of research in the oceanographic domain follows
that in spacecraft autonomy, from onboard \cite{mus98} to
mixed-initiative methods \cite{bresina05}, the challenges are
significantly different with the environment playing a far larger
role. Moreover, prediction and projection of a future course of action
for a robot given the level of uncertainty calls for representational
and architectural methods which would enable adaptation at different
levels of abstraction, from mission-planning to low level actuation
\kcomment{and control}.

Our efforts were initially targeted to full robotic autonomy; while
this is an important goal that we continue to strive towards, it is
increasingly clear that human-in-the-loop methods also have a sizable
role to play in ocean exploration and by extension to \kcomment{the
  field of} Marine Robotics. Not only is an embedded
\kcomment{deliberative agent} an important element of the effort
towards smarter vehicles, but we are working towards an implementation
of \rx which will perform multi-vehicle plan synthesis for the \ode.

A key role that \rx has played and will continue to play, is towards
\emph{event response} scenarios. In Section \ref{sec:results} we show
the role Machine Learning techniques have played thus far in in-situ
plan adaptation. This interaction between Planning and Machine
Learning, we believe, will play a significant role going forward in
exploration and discovery. Interpretation of incoming sensor data to
deal with scientific surprise and \kcomment{re-adaptation} for
opportunistic science will be critical to dealing with the persistent
problem of undersampling the ocean\footnote{Walter Munk of the Scripps
  Institute of Oceanography has famously stated \emph{``Most of the
    previous century could be called a “century of undersampling”''}
  -- Testimony to the U.S. Commission On Ocean Policy, 18 April
  2002}.

Much of what we know of the ocean was derived by exploring in the
blind. In the near future, we anticipate longer duration robotic
vehicles with enough onboard computational capacity to run
deliberative agents such as \rxe, make water-column measurements and
reliably track and characterize dynamic features. While our
\kcomment{own} experiments have already demonstrated such
capabilities, the exploration Vs. exploitation tradeoff is better
determined when vehicles can make choices over longer time periods
(weeks and months) to study the impacts of bio-geochemical interaction
driven by longer duration physical forcing \kcomment{and climate
  change}.

Given our poor understanding of the coastal ocean in particular,
mixed-initiative methods of command and control for Sampling with
heterogenous assets is an important goal that is clearly at our
doorstep. We believe deliberation is, but a small yet critical part of
the solution to unfolding open research problems for AI and Marine
Robotics.