### 2012-06-05 final-version/nddlshop.tex

\begin{figure}[h!]
\begin{verbatim}

// Locations (Home, SuperMarket, etc.)
class Location {
  string name;
  Location(string _name){
    name = _name;
  }
}

// Products (Milk, Banana, etc.) and
class Product {
  string name;
  Product(string _name) {
    name = _name;
  }
}

// ProductLocations (Banana can be found at SuperMarket, for example)
class ProductLocation {
  Location location;
  Product product;

  ProductLocation(Location _location, Product _product){
    location = _location;
    product = _product;
  }
}

// Use built-in Timeline functionality to enforce that an agent:
// a) Can't be at more than one place at a time.
// b) Can't Go more than one place at a time.
// c) Can't Go somewhere and be At somewhere at the same time.

class AgentLocation extends Timeline{
  predicate At {
    Location loc;
  }
  predicate Going {
    Location from;
    Location to;
  }
}
// In addition to having a location timeline, the agent can buy and own things.  
// Note that the actions and the location predicates can be concurrent 
// so they can't be on the same Timeline

class Agent {
  AgentLocation location;

  Agent() {
    location = new AgentLocation();
  }
  action Buy {
    Product product;
  }
  action Go {
    Location from;
    Location to;
  }
  predicate Own {
  	Product product;
  }
}

\end{verbatim}
\caption{\small \nd model of the Shopping Agent example.}
\label{fig:nddlshop}
\end{figure}